---
title: "bestPatterns.Rmd"
author: "Andreas Völlenklee"
date: "07/31/2014"
output: html_document
toc: yes
---

```{r echo=FALSE}
# some gobal definitions
library(digest) # used to calculate crc32
options(width=80)
path <- "/home/voellenk/Dropbox/constellation" # base path for project
```

## Synopsis 
In this document the performance of chart pattern among different groups of stocks is evaluated. It takes as input combinations-Rdata file(s). The best patterns are stored in a .Rdata file for further comparison among another subset of stocks / time horizons.

## Loading data
The time series are being analyzed by the Rscript `quantify_groups.R`. The performance of all possible pattern are calculated by the Rscript `calc_all_combinations.R`. As an output, either one or 25 .Rdata file with a list of data.frames `COMB.list` is stored in the Dropbox directory. This file / these files is read in with the following code chunk. 

```{r}
if (length(filenames == 1)) {
  load(paste(input.dir, filenames[1], sep="/"))
} else if (length(filenames == 25)) {
  C.list <- list()
  # read in all 25 files and combine them to one COMB.list
  for (filename in filenames) {
    load(paste(input.dir, filename, sep="/"))
    C.list[[names(COMB.list)]] <- COMB.list[[1]]
  }
  COMB.list <- C.list
} else {
  stop ("ERROR: Either one or 25 filenames must be given.")
}

# determine and print out key vector
key.vec <- c(1:(ncol(COMB.list[[1]])-2))
merge.cols <- colnames(COMB.list[[1]][key.vec])
merge.cols
```

```{r echo=FALSE}
# function to calculate a merged data.frame of defined groups
# groups is a vector of all groups that are taken for calculations
mergeCombs <- function(groups) {
  for (i in 1:length(groups)) {
    this.group <- groups[i]
    print(paste("doing group", this.group))
    this.df <- COMB.list[[this.group]]
    # in first iteration, prepare merged data.frame out of pattern columns
    if (i == 1) {
      merged.df <- this.df[,c(1:(ncol(this.df)-2))]
    }
    this.key <- this.df[,key.vec]
    this.performance <- cbind(this.df[,"buyS"], this.df[,"sellS"])[,c(1,3,4)] # omit redundand n column
    colnames(this.performance) <- c(paste(this.group, "med.buyS", sep="."),
                                    paste(this.group, "med.sellS", sep="."),
                                    paste(this.group, "n", sep="."))
    
    merged.df <- merge(merged.df, cbind(this.key, this.performance), by=merge.cols, all=TRUE)
  }
  return (merged.df)
}

# function to calculate summary.df
calcSummary <- function(merged.df, minimum.obs=20) {
  cols.buyS <- grep('med.buyS', names(merged.df))
  cols.sellS <- grep('med.sellS', names(merged.df))
  cols.n <- grep('.n', names(merged.df))
  
  summary.df <- merged.df[,key.vec]
  summary.df$mean.buyS <- apply(merged.df[,cols.buyS], 1, mean, na.rm=TRUE)
  summary.df$mean.sellS <- apply(merged.df[,cols.sellS], 1, mean, na.rm=TRUE)
  summary.df$mean.n <- apply(merged.df[,cols.n], 1, mean, na.rm=TRUE)
  summary.df$mean.n <- round(summary.df$mean.n)
  
  return (summary.df[summary.df$mean.n > minimum.obs, ])
}

# function adds column of key-ID to data.frame (previous method, no longer used)
addCRC32 <- function(rdf, key.vec) {
  df <- rdf[, key.vec]
  svec <- apply(df, 1, paste, collapse="")
  ID <- tapply(svec, 1:length(svec), digest, algo="crc32", ascii=TRUE)
  rownames(rdf) <- ID
  return(rdf)
}

# longer key to avoid multiple entries
addxxhash64 <- function(rdf, key.vec) {
  df <- rdf[, key.vec]
  svec <- apply(df, 1, paste, collapse="")
  ID <- tapply(svec, 1:length(svec), digest, algo="xxhash64", ascii=TRUE)
  rownames(rdf) <- ID
  return(rdf)
}

# simple version
addNames <- function(rdf, key.vec) {
  df <- rdf[, key.vec]
  
}
```

## Merging combinations of all groups
```{r echo=FALSE}
# all groups
groups <- names(COMB.list)
merged.df <- mergeCombs(groups)
merged.df <- addxxhash64(merged.df, key.vec)

# selection of groups
groups.low.sd <- groups[grep('^G_[12]', groups)]
merged.df.low.sd <- mergeCombs(groups.low.sd)
merged.df.low.sd <- addxxhash64(merged.df.low.sd, key.vec)

groups.high.sd <- groups[grep('^G_[345]', groups)]
merged.df.high.sd <- mergeCombs(groups.high.sd)
merged.df.high.sd <- addxxhash64(merged.df.high.sd, key.vec)

groups.low.ret <- groups[grep('[345]$', groups)]
merged.df.low.ret <- mergeCombs(groups.low.ret)
merged.df.low.ret <- addxxhash64(merged.df.low.ret, key.vec)

groups.high.ret <- groups[grep('[12]$', groups)]
merged.df.high.ret <- mergeCombs(groups.high.ret)
merged.df.high.ret <- addxxhash64(merged.df.high.ret, key.vec)

# quadrants Low sd, low return, etc.
groups.LSLR <- groups[grep('^G_[12]_[345]$', groups)]
merged.df.LSLR <- mergeCombs(groups.LSLR)
merged.df.LSLR <- addxxhash64(merged.df.LSLR, key.vec)

groups.LSHR <- groups[grep('^G_[12]_[12]$', groups)]
merged.df.LSHR <- mergeCombs(groups.LSHR)
merged.df.LSHR <- addxxhash64(merged.df.LSHR, key.vec)

groups.HSLR <- groups[grep('^G_[345]_[345]$', groups)]
merged.df.HSLR <- mergeCombs(groups.HSLR)
merged.df.HSLR <- addxxhash64(merged.df.HSLR, key.vec)

groups.HSHR <- groups[grep('^G_[345]_[12]$', groups)]
merged.df.HSHR <- mergeCombs(groups.HSHR)
merged.df.HSHR <- addxxhash64(merged.df.HSHR, key.vec)
```

The resulting data.frame merged.df contains the key columns and 3 columns for each group: `G_x_y.med.sellS`, `G_x_y.med.buyS` and `G_x_y.n`. The data.frame summary.df contains the key columns and 3 columns representing the mean value of `buyS`, `sellS`, `n`.

```{r}
summary.df <- calcSummary(merged.df)
summary.df.low.sd <- calcSummary(merged.df.low.sd)
summary.df.high.sd <- calcSummary(merged.df.high.sd)
summary.df.low.ret <- calcSummary(merged.df.low.ret)
summary.df.high.ret <- calcSummary(merged.df.high.ret)
summary.df.LSLR <- calcSummary(merged.df.LSLR)
summary.df.LSHR <- calcSummary(merged.df.LSHR)
summary.df.HSLR <- calcSummary(merged.df.HSLR)
summary.df.HSHR <- calcSummary(merged.df.HSHR)
```

```{r echo=FALSE}
# helper function: remove pattern with equal result. df must be sorted by top pattern descending.
reduceResult <- function(df, cutoff=10){
  n<-1
  i<-1
  while (n<=cutoff) {
    if (!(df[i,"mean.buyS"]==df[i+1,"mean.buyS"] && 
        df[i,"mean.sellS"]==df[i+1,"mean.sellS"] &&
        df[i,"mean.n"]==df[i+1,"mean.n"])) {
      this.row <- df[i,]
      if (n==1) {
        ret.df <- this.row
      } else {
        ret.df <- rbind(ret.df, this.row)
      }
      n <- n+1
    }
    i <- i+1    
  }
  return (ret.df)
}

consistentResults <- function(df, keycols, groups, min=.9) {
  perfcols <- c("mean.buyS", "mean.sellS", "mean.n")
  threshold <- ceiling(length(groups)*min)
  consistentBuy <- data.frame()
  consistentSell <- data.frame()
  for (i in 1:nrow(df)) {
    pos.groups <- 0
    neg.groups <- 0
    for (j in 1:length(groups)) {
      bc <- paste0(groups[j], ".med.buyS")
      sc <- paste0(groups[j], ".med.sellS")
      if (!is.na(df[i, bc]) && !is.na(df[i, sc])) {
        if (df[i, bc] >= df[i, sc]) {
          pos.groups <- pos.groups+1
        } else if (df[i, sc] >= df[i, bc]) {
          neg.groups <- neg.groups+1
        }        
      } else {
        pos.groups <-pos.groups+1
        neg.groups <-neg.groups+1
      }
    }
    if (pos.groups >= threshold) {
      consistentBuy <- rbind(consistentBuy, cbind(df[i,keycols], df[i,perfcols], pos.groups))
    } else if (neg.groups >= threshold) {
      consistentSell <- rbind(consistentSell, cbind(df[i,keycols], df[i,perfcols], neg.groups))
    }
  }
  return(list(consistentBuy, consistentSell))
}

### build list of best results out of a summary.df
bestResults <- function(summary.df, merged.df, groups, cutoff=300){
  res <- list()
  top.mean.buyS <- summary.df[with(summary.df, order(-mean.buyS)),]
  res[['topBuyS']] <- reduceResult(top.mean.buyS, cutoff=cutoff)
  
  top.mean.sellS <- summary.df[with(summary.df, order(-mean.sellS)),]
  res[['topSellS']] <- reduceResult(top.mean.sellS, cutoff=cutoff)

  top.max.pos.diff <- summary.df[with(summary.df, order(-(mean.buyS-mean.sellS))),]
  res[['topMaxDiffBuyS']] <- reduceResult(top.max.pos.diff, cutoff=cutoff)
  
  top.max.neg.diff <- summary.df[with(summary.df, order(-(mean.sellS-mean.buyS))),]
  res[['topMaxDiffSellS']] <- reduceResult(top.max.neg.diff, cutoff=cutoff)

  merged.top.df <- merge(rbind(res$topMaxDiffBuyS, res$topMaxDiffSellS), merged.df, by=merge.cols, all=FALSE)
  merged.top.df <- addxxhash64(merged.top.df, key.vec)
  cres.list <- consistentResults(merged.top.df, key.vec, groups, min=.8)
  
  if (nrow(cres.list[[1]]) > 0) {
    res[['consistentPos']] <- cres.list[[1]][with(cres.list[[1]], 
                           order(-pos.groups, -(mean.buyS-mean.sellS))), ][1:min(100,nrow(cres.list[[1]])),]    
  } else {
    res[['consistentPos']] <- data.frame()
  }

  # if statement to avoid execution error at sort command
  if (nrow(cres.list[[2]]) > 0) {
    res[['consistentNeg']] <- cres.list[[2]][with(cres.list[[2]], 
                           order(-neg.groups, -(mean.sellS-mean.buyS))), ][1:min(100,nrow(cres.list[[2]])),]   
  } else {
    res[['consistentNeg']] <- data.frame()    
  }

  return (res)
}
```

## Outstanding patterns
### overall all groups
#### highest mean signal
```{r}
res <- list()
res <- bestResults(summary.df, merged.df, groups)

res$topBuyS[1:16,]
res$topSellS[1:16,]
```

#### highest difference
```{r}
res$topMaxDiffBuyS[1:16,]
res$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res$consistentPos
res$consistentNeg
```


### groups with low sd
#### highest mean signal
```{r}
res.low.sd <- list()
res.low.sd <- bestResults(summary.df.low.sd, merged.df.low.sd, groups.low.sd)

res.low.sd$topBuyS[1:16,]
res.low.sd$topSellS[1:16,]
```

#### highest difference
```{r}
res.low.sd$topMaxDiffBuyS[1:16,]
res.low.sd$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.low.sd$consistentPos
res.low.sd$consistentNeg
```




### groups with high sd
#### highest mean signal
```{r}
res.high.sd <- list()
res.high.sd <- bestResults(summary.df.high.sd, merged.df.high.sd, groups.high.sd)

res.high.sd$topBuyS[1:16,]
res.high.sd$topSellS[1:16,]
```

#### highest difference
```{r}
res.high.sd$topMaxDiffBuyS[1:16,]
res.high.sd$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.high.sd$consistentPos
res.high.sd$consistentNeg
```




### groups with low median return
```{r}
res.low.ret <- list()
res.low.ret <- bestResults(summary.df.low.ret, merged.df.low.ret, groups.low.ret)

res.low.ret$topBuyS[1:16,]
res.low.ret$topSellS[1:16,]
```

#### highest difference
```{r}
res.low.ret$topMaxDiffBuyS[1:16,]
res.low.ret$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.low.ret$consistentPos
res.low.ret$consistentNeg
```




### groups with high median return
```{r}
res.high.ret <- list()
res.high.ret <- bestResults(summary.df.high.ret, merged.df.high.ret, groups.high.ret)

res.high.ret$topBuyS[1:16,]
res.high.ret$topSellS[1:16,]
```

#### highest difference
```{r}
res.high.ret$topMaxDiffBuyS[1:16,]
res.high.ret$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.high.ret$consistentPos
res.high.ret$consistentNeg
```




### groups in quadrant low sd, low return
```{r}
res.LSLR <- list()
res.LSLR <- bestResults(summary.df.LSLR, merged.df.LSLR, groups.LSLR)

res.LSLR$topBuyS[1:16,]
res.LSLR$topSellS[1:16,]
```

#### highest difference
```{r}
res.LSLR$topMaxDiffBuyS[1:16,]
res.LSLR$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.LSLR$consistentPos
res.LSLR$consistentNeg
```




### groups in quadrant low sd, high return
```{r}
res.LSHR <- list()
res.LSHR <- bestResults(summary.df.LSHR, merged.df.LSHR, groups.LSHR)

res.LSHR$topBuyS[1:16,]
res.LSHR$topSellS[1:16,]
```

#### highest difference
```{r}
res.LSHR$topMaxDiffBuyS[1:16,]
res.LSHR$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.LSHR$consistentPos
res.LSHR$consistentNeg
```




### groups in quadrant high sd, low return
```{r}
res.HSLR <- list()
res.HSLR <- bestResults(summary.df.HSLR, merged.df.HSLR, groups.HSLR)

res.HSLR$topBuyS[1:16,]
res.HSLR$topSellS[1:16,]
```

#### highest difference
```{r}
res.HSLR$topMaxDiffBuyS[1:16,]
res.HSLR$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.HSLR$consistentPos
res.HSLR$consistentNeg
```




### groups in quadrant high sd, high return
```{r}
res.HSHR <- list()
res.HSHR <- bestResults(summary.df.HSHR, merged.df.HSHR, groups.HSHR)

res.HSHR$topBuyS[1:16,]
res.HSHR$topSellS[1:16,]
```

#### highest difference
```{r}
res.HSHR$topMaxDiffBuyS[1:16,]
res.HSHR$topMaxDiffSellS[1:16,]
```

#### pattern with consistent buyS > sellS in all groups
```{r}
res.HSHR$consistentPos
res.HSHR$consistentNeg
```

## Store results
The results of this report are stored in a .RData variable in the Dropbox directory /constellation/bestPatterns. 
```{r}
savepath <- paste(path, "bestPatterns", sep="/")
savefilename <- sub("help=FALSE_C.Rdata$", "bestPatterns.Rdata", filename)
save(list=c("res", "res.low.sd", "res.high.sd", "res.low.ret", "res.high.ret",
            "res.LSLR", "res.LSHR", "res.HSLR", "res.HSHR"), 
     file=paste(savepath, savefilename, sep="/"))
```